---
title: "Using CollaborativeBookdown"
author:
- name: Sean Davis
  affiliation: National Cancer Institute, National Institutes of Health
  email: seandavi@gmail.com
package: CollaborativeBookdown
output:
  BiocStyle::html_document
abstract: |
  Description of your vignette
vignette: |
  %\VignetteIndexEntry{Using CollaborativeBookdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

The `r CRANpkg("bookdown")` package can be used to build gitbook books from
multiple R markdown documents. The process for doing so is well-documented for
a single user with a directory of R markdown documents. 

A more complicated use case includes some additional requirements.

- The materials and R packages required to successfully run the materials
in the book should be installable, including dependencies.
- Book chapters will be authored by individual contributors and each should
be able to control when new content is merged into the book while still
being able to work independently on their own materials.
- Contributions from each user should themselves be versioned, coupled to a
particular version of the book, and potentially decoupled from the book to 
include or exclude content.
- Individual chapters and (and should) be incorporated into individual installable
R packages so that R dependencies, code, and extended documentation can be included.
- The bookdown "parent" repository will depend on the packages from individual 
contributors. Installing the "parent" repository will lead to versioned installations
of each "chapter" repository.


# Introduction

```{r message=FALSE, warning=FALSE}
library(CollaborativeBookdown)
```

# Usage

## Testing contributed packages

The first concern when building a collaborative bookdown
book based on workshop packages is to test which of the 
packages actually installs. The `testPackageInstalls()` function
takes a character vector of [repository strings] **PUT IN REMOTES HERE**.
The convenience function `repoFromURLs()` can convert from plain
URLs to the format needed for installation.


```{r testpkginstall, message=FALSE, warning=FALSE, eval=FALSE}
z1 = testPackageInstalls(repos)
head(z1)
```

Once testing for individual package installations has completed,
a second step is to test and install the package **including 
vignette build and installation**. The output will again be a
three-column data.frame with details of the install process.

Calling the `testPackageInstalls()` function again, but with additional
parameters will install each package, build vignettes, and install those
as well.


```{r message=FALSE, warning=FALSE, eval=FALSE}
z2 = testPackageInstalls(repos, build=TRUE, build_opts="", force=TRUE)
```

# Package vignettes to bookdown

## Finding installed package vignettes

Since the `r Githubpkg('seandavi/CollaborativeBookdown')` package needs
to have access to the *R markdown* source which should have been installed
with the package, a convenience function returns a data frame of packages, 
absolute paths to each vignette, and the vignette filename (for easy linking
if that is desirable).

```{r installedVignettes}
ivignettes = installedVignettes(c('BiocManager','GEOquery', 'dplyr', 'abc'))
ivignettes
```


## Checking package vignettes


```{r fakeRmdFile}
temp = tempfile()
rmdtest = paste("# A test of multiple headers",
"", "# second header", "",
"And some text", sep="\n")
writeLines(rmdtest, temp)
```

Bookdown has some special requirements for R markdown files that are going
to appear as chapters in a bookdown book.

1. Chapters must contain one and only one chapter header.

The `checkVignette()` function performs some basic checks to see if the 
supplied `.Rmd` files meet these criteria.

```{r}
checkVignette(temp)
```

## Fixing common vignette problems

The test `.Rmd` file has incorrectly specified the number of headers. The
`fixMultipleFirstLevelHeaders()` function implements a kludge to fix the file.

```{r}
cat(paste(fixMultipleFirstLevelHeaders(temp),collapse="\n"))
```

The new text has demoted all the first-level headers--except the first one--to 
second-level headers. If necessary, use `writeLines` to write out the results
to a file.



# sessionInfo()

```{r}
devtools::session_info()
```