---
title: "Using CollaborativeBookdown"
author:
- name: Sean Davis
  affiliation: National Cancer Institute, National Institutes of Health
  email: seandavi@gmail.com
package: CollaborativeBookdown
output:
  BiocStyle::html_document
abstract: |
  Description of your vignette
vignette: |
  %\VignetteIndexEntry{Using CollaborativeBookdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

The `r CRANpkg("bookdown")` package can be used to build gitbook books
from multiple R markdown documents. The process for doing so is
well-documented for a single user with a directory of R markdown
documents.

A more complicated use case includes some additional requirements.

- The materials and R packages required to successfully run the
materials in the book should be installable, including dependencies.
- Book chapters will be authored by individual contributors and each
should be able to control when new content is merged into the book
while still being able to work independently on their own materials.
- Contributions from each user should themselves be versioned, coupled
to a particular version of the book, and potentially decoupled from
the book to include or exclude content.
- Individual chapters and (and should) be incorporated into individual
installable R packages so that R dependencies, code, and extended
documentation can be included.
- The bookdown "parent" repository will depend on the packages from
individual contributors. Installing the "parent" repository will lead
to versioned installations of each "chapter" repository.

## Implementation assumptions

- Each chapter will be derived from a vignette from an installable
  package.
- Each resource package that includes one or more chapter vignettes
will be a self-contained and installable.
- Each vignette in each resource package will use `system.file()` to
access assets such as images, bibliography, etc. to enable compiling
the vignette *after* installation.

## Additional needs not yet discussed or implemented

- Since a package may include multiple vignettes, consider adding an
  .workshop_config file to each package in `inst` that will include
  details of which vignette to include.
- Given a set of workshop packages, there should be an approach to
  reproducibly and reliably build and test each package through to
  complete artifact creation (gitbook, etc.)
- The build environment should be reproducible and available to
  workshop authors so that authors can be assured of a successful
  build when combined with other workshop chapters.
- Build materials for various vignettes should be isolated from each
  other. **Not sure how best to go about this**. The issue is that
  when building multiple chapters together, gitbook needs `.Rmd` files
  in one location, so need to ensure that chapters are "namespaced"
  somehow. 
- Consider a workshop "manifest" that includes remotes, git hashes (to
  pin versions), and perhaps even a packrat repo to pin package
  versions.



# Usage

```{r message=FALSE, warning=FALSE}
library(CollaborativeBookdown)
```

## Testing contributed packages

The first concern when building a collaborative bookdown book based on
workshop packages is to test which of the packages actually
installs. The `testPackageInstalls()` function takes a character
vector of [repository strings] **PUT IN REMOTES HERE**.  The
convenience function `repoFromURLs()` can convert from plain URLs to
the format needed for installation.


```{r testpkginstall, message=FALSE, warning=FALSE, eval=FALSE}
z1 = testPackageInstalls(repos)
head(z1)
```

Once testing for individual package installations has completed,
a second step is to test and install the package **including 
vignette build and installation**. The output will again be a
three-column data.frame with details of the install process.

Calling the `testPackageInstalls()` function again, but with additional
parameters will install each package, build vignettes, and install those
as well.


```{r message=FALSE, warning=FALSE, eval=FALSE}
z2 = testPackageInstalls(repos, build=TRUE, build_opts="", force=TRUE)
```

# Package vignettes to bookdown

## Finding installed package vignettes

Since the `r Githubpkg('seandavi/CollaborativeBookdown')` package needs
to have access to the *R markdown* source which should have been installed
with the package, a convenience function returns a data frame of packages, 
absolute paths to each vignette, and the vignette filename (for easy linking
if that is desirable).

```{r installedVignettes}
ivignettes = installedVignettes(c('BiocManager','GEOquery', 'dplyr', 'abc'))
ivignettes
```


## Checking package vignettes

Create a fake vignette as text and write to file for testing and
demonstrating functionality.

```{r fakeRmdFile}
temp = tempfile()
rmdtest = paste("# A test of multiple headers",
"", "# second header", "",
"And some text", sep="\n")
writeLines(rmdtest, temp)
```

Bookdown has some special requirements for R markdown files that are going
to appear as chapters in a bookdown book.

1. Chapters must contain one and only one chapter header (first-level
   heading).

The `checkVignette()` function performs some basic checks to see if the 
supplied `.Rmd` files meet these criteria.

```{r}
checkVignette(temp)
```

## Fixing common vignette problems

The test `.Rmd` file has incorrectly used more than one first-level
header, the number of headers. The `fixMultipleFirstLevelHeaders()`
function implements a kludge to fix the file.

```{r}
cat(paste(fixMultipleFirstLevelHeaders(temp),collapse="\n"))
```

The new text has demoted all the first-level headers--except the first
one--to second-level headers. If necessary, use `writeLines` to write
out the results to a file.


# sessionInfo()

```{r}
devtools::session_info()
```
